---
title: "DeadWood_Biomass"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DeadWood_Biomass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

`DeadWood_Biomass` translates snag and DWD decay class inventories into
pixel-level biomass estimates (Mg ha⁻¹) by joining each record to a
species- and pool-specific density reduction factor (DRF) lookup table.
Optionally saves multi-layer biomass snapshot rasters at a configurable
plot interval.

## Inputs

| Object | Class | Description |
|--------|-------|-------------|
| `snagTable` | `data.table` | Current snag inventory from `DeadWood_snagDecay` |
| `DWDTable` | `data.table` | Current DWD inventory from `DeadWood_DWDDecay` |
| `studyAreaRaster` | `SpatRaster` | Template raster defining pixel grid, CRS, and resolution |

## Outputs

| Object | Class | Description |
|--------|-------|-------------|
| `snagBiomass_Mg_ha` | `SpatRaster` | Pixel-level snag biomass (Mg ha⁻¹) |
| `DWDBiomass_Mg_ha` | `SpatRaster` | Pixel-level DWD biomass (Mg ha⁻¹) |
| `snagHistory` | `SpatRaster` | Multi-layer snag biomass snapshots (one layer per plot interval) |
| `DWDHistory` | `SpatRaster` | Multi-layer DWD biomass snapshots (one layer per plot interval) |

## Parameters

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `DRFLookup` | `data.table` | Empty table | Density reduction factors: `species`, `pool`, `DC`, `DRF` |
| `.plotInitialTime` | `numeric` | `NA` | Simulation time for first snapshot. `NA` = no snapshots |
| `.plotInterval` | `numeric` | `1` | Years between snapshots |

## Example

```{r example}
library(SpaDES.core)
library(data.table)
library(terra)

DRFLookup <- data.table(
  species = "Pinus strobus",
  pool    = rep(c("snag", "DWD"), each = 5),
  DC      = rep(1:5, 2),
  DRF     = c(1.000, 0.841, 0.706, 0.543, 0.382,
              1.000, 0.783, 0.614, 0.418, 0.251)
)

mySim <- simInit(
  times   = list(start = 0, end = 50),
  modules = list("DeadWood_Biomass"),
  params  = list(DeadWood_Biomass = list(
    DRFLookup        = DRFLookup,
    .plotInitialTime = 5,
    .plotInterval    = 5
  )),
  objects = list(
    snagTable       = snagTable,
    DWDTable        = DWDTable,
    studyAreaRaster = myRaster
  ),
  paths = list(modulePath = "path/to/modules")
)

mySim <- spades(mySim)
```
