---
title: "DeadWood_Biomass"
author: ""
#date: "`r Sys.Date()`"
output: github_document
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_file = "README.md",
                      output_format = "github_document") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

`DeadWood_Biomass` translates snag and DWD decay class inventories into
pixel-level biomass estimates (Mg ha⁻¹) by summing initial biomass at death
(`initBiomass`) across all pieces within each pixel. Optionally saves
multi-layer biomass snapshot rasters at a configurable plot interval.

## Inputs

| Object | Class | Description |
|--------|-------|-------------|
| `snagTable` | `data.table` | Current snag inventory from `DeadWood_snagDecay`. Must have columns `pixelID` (integer), `initBiomass` (numeric, Mg ha⁻¹), and `species` (character). |
| `DWDTable` | `data.table` | Current DWD inventory from `DeadWood_DWDDecay`. Must have columns `pixelID` (integer), `initBiomass` (numeric, Mg ha⁻¹), and `species` (character). |
| `studyAreaRaster` | `SpatRaster` | Template raster defining the spatial grid (extent, CRS, resolution). All output rasters are created to match this template. |

## Outputs

| Object | Class | Description |
|--------|-------|-------------|
| `snagBiomass_Mg_ha` | `SpatRaster` | Pixel-level total snag biomass (Mg ha⁻¹), updated every 5 years. Pixels with no snags have value `NA`. |
| `DWDBiomass_Mg_ha` | `SpatRaster` | Pixel-level total DWD biomass (Mg ha⁻¹), updated every 5 years. Pixels with no DWD have value `NA`. |
| `snagHistory` | `SpatRaster` | Multi-layer raster accumulating one snag biomass snapshot per plot event, all species combined (layers named `yr<time>`). `NULL` if `.plotInitialTime` is `NA`. |
| `DWDHistory` | `SpatRaster` | Multi-layer raster accumulating one DWD biomass snapshot per plot event, all species combined (layers named `yr<time>`). `NULL` if `.plotInitialTime` is `NA`. |
| `snagHistoryBySpecies` | `list` | Named list of multi-layer SpatRasters of snag biomass snapshots, one entry per species. |
| `DWDHistoryBySpecies` | `list` | Named list of multi-layer SpatRasters of DWD biomass snapshots, one entry per species. |

## Parameters

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `.plotInitialTime` | `numeric` | `NA` | Simulation time of the first snapshot. Set to `NA` to disable snapshot accumulation entirely. |
| `.plotInterval` | `numeric` | `5` | Interval (years) between snapshots. Defaults to 5 to align with the decay timestep. Only used if `.plotInitialTime` is not `NA`. |

## Events

### `init` (once, at simulation start)

- Creates `snagBiomass_Mg_ha` and `DWDBiomass_Mg_ha` as copies of
  `studyAreaRaster` filled with `NA`.
- Initialises `snagHistory`, `DWDHistory` to `NULL` and
  `snagHistoryBySpecies`, `DWDHistoryBySpecies` to empty lists.
- Schedules the first `transition` event at `start(sim) + 5` with priority 4.
- If `.plotInitialTime` is not `NA`, schedules the first `plot` event at
  `.plotInitialTime` with priority 5.

### `transition` (every 5 years, priority 4)

Runs after `DeadWood_snagDecay` and `DeadWood_DWDDecay` have updated their
inventories. For each pool, `initBiomass` (biomass at time of death) is summed
by pixel:

$$\text{snagBiomass}[p] = \sum_{i \,:\, \text{pixelID}_i = p} \text{initBiomass}_i$$

$$\text{DWDBiomass}[p] = \sum_{i \,:\, \text{pixelID}_i = p} \text{initBiomass}_i$$

where $p$ is a pixel index and $i$ indexes individual pieces. Pixels with no
pieces receive `NA`. Note: `initBiomass` represents biomass at time of death
and has not been adjusted by a density reduction factor (DRF). DRF application
is not yet implemented.

### `plot` (optional, every `.plotInterval` years, priority 5)

Captures snapshots of `snagBiomass_Mg_ha` and `DWDBiomass_Mg_ha` at the
current simulation time and appends them as named layers to `snagHistory` and
`DWDHistory`. Also appends per-species layers to `snagHistoryBySpecies` and
`DWDHistoryBySpecies`. Layer names follow the pattern `yr<time>`
(e.g., `yr10`, `yr25`). Snapshot events continue until the end of the
simulation.

## Event scheduling and module interactions

This module runs at the lowest priority within each 5-year timestep, after both
decay modules have updated their inventories. `DeadWood_Mortality` fires every
year; the remaining modules fire every 5 years.

| Timestep | Priority | Module | Event | Purpose |
|----------|----------|--------|-------|---------|
| every 1 yr | 0 | `DeadWood_Mortality` | `transition` | Apply baseline mortality, append to `cohortData` |
| every 5 yr | 1 | `DeadWood_snagDecay` | `transition` | Advance snag DC, produce `fallenSnags` |
| every 5 yr | 2 | `DeadWood_DWDDecay` | `receive` | Accept `fallenSnags` into DWD pool |
| every 5 yr | 3 | `DeadWood_DWDDecay` | `transition` | Advance DWD DC via logistic model |
| every 5 yr | 4 | `DeadWood_Biomass` | `transition` | Compute biomass rasters ← **this module** |
| every 5 yr | 5 | `DeadWood_Biomass` | `plot` | Capture biomass snapshots ← **this module** |

## Example

```{r example}
library(SpaDES.project)

# Simulate a post-disturbance gradient across a 5x5 landscape:
# snag biomass is high near the disturbance origin (top rows, pixels 1-10)
# and declines with distance. DWD shows the inverse pattern — older fallen
# material has accumulated in the lower rows over time.
snagTable <- data.table::data.table(
  pixelID     = 1:25,
  species     = c(rep("Pinus strobus", 13), rep("Pinus resinosa", 12)),
  DC          = rep(1:5, 5),
  ageInDC     = 0L,
  initBiomass = c(seq(30, 8, length.out = 13), seq(22, 6, length.out = 12)),
  diameter_cm = 25.0
)

DWDTable <- data.table::data.table(
  pixelID       = 1:25,
  species       = c(rep("Pinus strobus", 13), rep("Pinus resinosa", 12)),
  DC            = rep(1:5, 5),
  ageInDC       = 0L,
  ageSinceEntry = 0L,
  initBiomass   = c(seq(8, 25, length.out = 13), seq(5, 18, length.out = 12)),
  diameter_cm   = 25.0
)

studyAreaRaster <- terra::rast(nrows = 5, ncols = 5, vals = 0)

out <- SpaDES.project::setupProject(
  paths           = list(modulePath = "modules"),
  modules         = "BosunForestEcology/DeadWood_Biomass@main",
  times           = list(start = 0, end = 10),
  snagTable       = snagTable,
  DWDTable        = DWDTable,
  studyAreaRaster = studyAreaRaster
)

mySim <- SpaDES.core::simInitAndSpades2(out)

par(mfrow = c(1, 2))
terra::plot(mySim$snagBiomass_Mg_ha, main = "Snag biomass (Mg/ha)")
terra::plot(mySim$DWDBiomass_Mg_ha,  main = "DWD biomass (Mg/ha)")
par(mfrow = c(1, 1))
```

## Package dependencies

- [`SpaDES.core`](https://github.com/PredictiveEcology/SpaDES.core) (>= 3.0.0)
- [`data.table`](https://CRAN.R-project.org/package=data.table)
- [`terra`](https://CRAN.R-project.org/package=terra)
- [`ggplot2`](https://CRAN.R-project.org/package=ggplot2)
